---
title: "Startup Dataset Description"
output: html_notebook
---
<br>
    <h3>The Goal of Collecting this Dataset</h3>
  
Our <b>goal</b> of collecting this dataset is to build a model capable of predicting 
the potential success of startups.

<br>
    <h3>The Source of the Dataset</h3>

Access the dataset on Kaggle through the following link: https://www.kaggle.com/datasets/manishkc06/startup-success-prediction

<br>
    <h3>General Information about the Dataset</h3>

Number of objects: <b>924 rows</b>.

Number of attributes: <b>20 columns</b>.

In this dataset, the <b>class label</b> we are focusing on is "<b>status</b>", which distinguishes between 
<b>successful</b> startups (labeled as "<b>acquired</b>") and <b>unsuccessful</b> ones (labeled as "<b>closed</b>").

**************************************************************************************************

```{r}
# Import the readr library
library(readr)
```

```{r}
# Specify the dataset path
dataset_path <- "/Users/basmaalsulaim/Desktop/IT326 Project/Dataset/StartupData.csv"
```

```{r}
# Load the dataset
dataset <- read_csv(dataset_path)
```
```{r}
# View the first few rows of the dataset
head(dataset)
```
```{r}
# Summarize all data
summary(dataset)
```