---
title: "Startup Dataset Description"
output: html_notebook
---
### The Goal of Collecting this Dataset
  
Our **goal** of collecting this dataset is to build a model capable of predicting the potential success of startups.

### The Source of the Dataset

Access the dataset on Kaggle through the following link: https://www.kaggle.com/datasets/manishkc06/startup-success-prediction

### General Information about the Dataset

**Number of attributes:** 50 column
**Number of objects:** 925 row
**Class label:** status ("acquired" --> successful company, or "closed" --> unsuccessful company)
**Types of attributes:** Nominal, Numerical, Binary

| Attribute Name           | Description                                                  | Data Type | Possible Values                                 |
| ------------------------ | ------------------------------------------------------------ | --------- | ----------------------------------------------- |
| Unnamed: 0               | Method of numbering companies.                               | Nominal   | 1 to 1153                                       |
| state_code               | The code of the state the startup was founded in.            | Nominal   | Different state codes. Like CA, AZ              |
| latitude                 | The latitude of the startup headquarters.                    | Numerical | 90 to -90                                       |
| longitude                | The longitude of the startup headquarters.                   | Numerical | 0 to 180                                        |
| zip_code                 | The zip code of the city the startup was founded in.         | Nominal   | Different random city zipcodes                  |
| id                       | N/A                                                          | Nominal   | Data in unrecognizable format c:number          |
| city                     | The name of the city the startup was founded in.             | Nominal   | Different cities. Like Palo Alto, Mountain View |
| Unnamed: 6               | The address of the headquarters.                             | Nominal   | Different addresses. Like San Diego CA 92121    |
| name                     | The name of the startup.                                     | Nominal   | Different names. Like Qsecure, drop.io          |
| labels                   | N/A                                                          | Binary    | 0 or 1                                          |
| founded_at               | Startup founding date.                                       | Nominal | 01 1984 to 09 2010                              |
| closed_at                | Startup closing date.                                        | Nominal | 01 2001 to 08 2013                              |
| first_funding_at         | Startup first funding date.                                  | Nominal | 01 2000 to 09 2009                              |
| last_funding_at          | Startup last funding date.                                   | Nominal | 01 2001 to 09 2011                              |
| age_first_funding_year   | The average age of startup when received first funding.      | Numerical | 0 to 21                                         |
| age_last_funding_year    | The average age of startup when received last funding.       | Numerical | 0 to 21                                         |
| age_first_milestone_year | The average age of startup when acheived first milestone.    | Numerical | 0 to 24                                         |
| age_last_milestone_year  | The average age of startup when acheived last milestone.     | Numerical | 0 to 24                                         |
| relationships            | N/A                                                          | Numerical | 0 to 63                                         |
| funding_rounds           | The number of funding rounds the startup went through.       | Numerical | 1 to 10                                         |
| funding_total_usd        | The total number of money in USD raised by the startup.      | Numerical | 11000 to 5700000000                             |
| milestones               | The total number of milestones achieved by the startup.      | Numerical | 0 to 8                                          |
| state_code.1             | The code of the state the startup was founded in.            | Nominal   | Different state codes. Like CA, AZ              |
| is_CA                    | If the startup was founded in California.                    | Binary    | 0 or 1                                          |
| is_NY                    | If the startup was founded in New York.                      | Binary    | 0 or 1                                          |
| is_MA                    | If the startup was founded in Massachusetts.                 | Binary    | 0 or 1                                          |
| is_TX                    | If the startup was founded in Texas.                         | Binary    | 0 or 1                                          |
| is_otherstate            | If the startup was founded in a state other than the listed. | Binary    | 0 or 1                                          |
| category_code            | The sector of the startup.                                   | Nominal   | Different categories. Like biotech, mobile   |
| is_software              | If the startup sector is software.                           | Binary    | 0 or 1                                          |
| is_web                   | If the startup sector is web.                                | Binary    | 0 or 1                                          |
| is_mobile                | If the startup sector is mobile.                             | Binary    | 0 or 1                                          |
| is_enterprise            | If the startup sector is enterprise.                         | Binary    | 0 or 1                                          |
| is_advertising           | If the startup sector is advertising.                        | Binary    | 0 or 1                                          |
| is_gamesvideo            | If the startup sector is video games.                        | Binary    | 0 or 1                                          |
| is_ecommerce             | If the startup sector is ecommerce.                          | Binary    | 0 or 1                                          |
| is_biotech               | If the startup sector is biotech.                            | Binary    | 0 or 1                                          |
| is_consulting            | If the startup sector is consulting.                         | Binary    | 0 or 1                                          |
| is_othercategory         | If the startup sector is any other category that the listed. | Binary    | 0 or 1                                          |
| object_id                | N/A                                                          | Nominal   | Data in unrecognizable format c:number          |
| has_VC                   | If the startup has a venture capitalist\* investor.          | Binary    | 0 or 1                                          |
| has_angel                | If the startup has an angel\*\* investor.                    | Binary    | 0 or 1                                          |
| has_roundA               | If the startup went through a series A funding round.        | Binary    | 0 or 1                                          |
| has_roundB               | If the startup went through a series B funding round.        | Binary    | 0 or 1                                          |
| has_roundC               | If the startup went through a series C funding round.        | Binary    | 0 or 1                                          |
| has_roundD               | If the startup went through a series D funding round.        | Binary    | 0 or 1                                          |
| avg_participants         | The average number of participants in the startup.           | Binary    | 0 or 1                                          |
| is_top500                | If the startup is a top 500 company.                         | Binary    | 0 or 1                                          |
| status                   | The acquisition status of the startup.                       | Binary    | "acquired" or "closed"                          |

***VC** stands for Venture Capitalist. Which is a type of investor that invests the money of a venture capital firm into small startup companies.

\*\***Angel** stands for Angel Investor Which is a type of investor that invests their personal money  into small startup companies in exchange for a percentage in the company.

```{r}
file_path <- "Original_StartupData.csv"
OriginalStartupData <- read.csv(file_path)
```

```{r}
# View the first few rows of the dataset
head(OriginalStartupData)
```

```{r}
# View the last few rows of the dataset
tail(OriginalStartupData)
```

```{r}
# Summarize all data
summary(OriginalStartupData)
```